Analytic reproducibility in articles receiving open data badges at Psychological Science: An observational assessment

### 2020-06-24 manuscript/manuscript.Rmd
shifting order of preprocessing and loading functions to avoid accidental removal of functions
affiliation changes and inclusion of meta-analysis

author: 
  - name          : "Tom E. Hardwicke"
    affiliation   : "1,2"
    corresponding : yes    # Define only one corresponding author
    address       : "Nieuwe Achtergracht 129B, Department of Psychology, University of Amsterdam, 1018 WT Amsterdam, The Netherlands"
    email         : "tom.hardwicke@uva.nl"
  - name          : "Manuel Bohn"
    affiliation   : "3"
  - name          : "Kyle MacDonald"
    affiliation   : "4"
  - name          : "Emily Hembacher"
    affiliation   : "3"    
  - name          : "Michèle B. Nuijten"
    affiliation   : "5"
  - name          : "Benjamin N. Peloquin"
    affiliation   : "3"
  - name          : "Benjamin E. deMayo"
    affiliation   : "3"
  - name          : "Bria Long"
    affiliation   : "3"
  - name          : "Erica J. Yoon"
    affiliation   : "3"
  - name          : "Michael C. Frank"
    affiliation   : "3"

affiliation:
  - id            : "1"
    institution   : "University of Amsterdam"
  - id            : "2"
    institution   : "Charité – Universitätsmedizin Berlin"
  - id            : "3"
    institution   : "Stanford University"
  - id            : "4"
    institution   : "University of California, Los Angeles"
  - id            : "5"
    institution   : "Tilburg University"

@@

abstract: | 
  For any scientific report, repeating the original analyses upon the original data should yield the original outcomes. We evaluated analytic reproducibility in `r knitr::load_cache(label = 'abstract_text', object = 'totalArticles')` Psychological Science articles awarded open data badges between 2014-2015. Initially, `r knitr::load_cache(label = 'abstract_text', object = 'intialReproRateArticles')` articles contained at least one “major numerical discrepancy” (>10% difference) prompting us to request input from original authors. Ultimately, target values were reproducible without author involvement for `r knitr::load_cache(label = 'abstract_text', object = 'repro_noAI')` articles; reproducible with author involvement for `r knitr::load_cache(label = 'abstract_text', object = 'repro_withAI')` articles; not fully reproducible with no substantive author response for `r knitr::load_cache(label = 'abstract_text', object = 'noRepro_noAI')` articles; and not fully reproducible despite author involvement for `r knitr::load_cache(label = 'abstract_text', object = 'noRepro_withAI')` articles. Overall, `r knitr::load_cache(label = 'abstract_text', object = 'majorErrors')` major numerical discrepancies remained out of `r knitr::load_cache(label = 'abstract_text', object = 'totalValues')` checked values `r knitr::load_cache(label = 'abstract_text', object = 'reproRateValues')`, but original conclusions did not appear affected. Non-reproducibility was primarily caused by unclear analysis reporting. Open data alone is not sufficient to ensure analytic reproducibility.

@@

```{r load_packages}
# load packages
library(knitr) # for literate programming
library(papaja) # for article template
library(tidyverse) # for data munging
library(tidylog) # inline analysis feedback
library(here) # for finding files
library(DescTools) # for calculating confidence intervals
library(english) # for converting numbers to words
```

@@

```{r load_data}
# loads the processed data files output by the preprocessing performed above
load(here('data', 'processed', 'data_values.RData')) # load data
load(here('data', 'processed', 'data_articles.RData')) # load data
load(here('data', 'processed', 'data_kidwell.RData')) # load data
```

@@

[^compRepro]: Assessments of computational reproducibility typically attempt to reproduce values by re-running original computational code with the original data and can therefore fail if original code is unavailable or non-functioning [e.g., @stodden2018]. By contrast, assessments of analytic reproducibility typically attempt to reproduce values by repeating the original analysis procedures, which can involve implementing those procedures in new code [e.g., @hardwicke2018a].

@@

Of `r data_kidwell_badge %>% nrow()` articles marked with an open data badge, Kidwell and colleagues had identified `r data_kidwell_badge_reusable %>% nrow()` with datasets that met four reusability criteria (accessible, correct, complete, and understandable). For each of these articles, one investigator (TEH) attempted to identify a coherent set of descriptive and inferential statistics, roughly 2-3 paragraphs of text, sometimes including a table or figure, related to a ‘substantive’ finding based on ‘relatively straightforward’ analyses. We focused on substantive findings because they are the most important and straightforward analyses to ensure that our team had sufficient expertise to re-run them. In total, `r nrow(data_values)` discrete numerical values reported in `r totalArticles` articles were identified as target values. The articles were published between January 2014 and May 2015. Further information about the sample is available in Supplementary Information \@ref(sup_sample).

@@

@@

```{r}
# prepare information on insufficent information errors for reporting inline
insuffInfoErrors <- data_articles %>% filter(Insufficient_Information_Errors > 0) %>% nrow() # how many articles had at least 1 insufficient information error
insuffInfoErrors_percent <- insuffInfoErrors/totalArticles*100 # calculate percent
insuffInfoErrors_CIs <- propCI(insuffInfoErrors,totalArticles) # caluclate CIs
```
Prior to author involvement, our reproducibility checks indicated that `r reproBefore %>% filter(reproOutcome == F) %>% pull(n)` out of the `r totalArticles` (`r reproBefore %>% filter(reproOutcome == F) %>% pull(percent)`) articles contained at least one major numerical discrepancy. After requesting input from original authors, several issues were resolved through the provision of additional information[^authorInfo] leaving `r reproAfter %>% filter(reproOutcome == F) %>% pull(n)` out of the `r totalArticles` (`r reproAfter %>% filter(reproOutcome == F) %>% pull(percent)`) articles containing at least one major numerical discrepancy. The full breakdown of reproducibility outcomes after author contact are shown in Table \@ref(tab:reproTab). Aspects of `r as.english(insuffInfoErrors)` reproducibility checks could not be completed as there was insufficient information to proceed with the reanalysis. Availability of original analysis scripts appeared to provide some modest benefits in terms of reproducibility outcomes and time expenditure (see Supplementary Information \@ref(sup_results)). In no cases did the observed numerical discrepancies appear to be consequential for the conclusions stated in the original articles (see Supplementary Information \@ref(vignettes)).

@@

```{r scatterPlot, fig.cap="Scatterplot showing reanalysis p-values as a function of reported p-values, classified by reproducibility outcome. Axes are on a log scale. For display purposes, 41 values below 0.001 are not shown.", fig.path="figs/"}
source(here('analysis','plot_reproScatter.R')) # build plot
scatterPlot # show plot
```

@@

A reasonable expectation of any scientific manuscript is that repeating the original analyses upon the original data will yield the same quantitative outcomes [@bollen2015]. We have found that this standard of analytic reproducibility was frequently not met in a sample of Psychological Science articles receiving open data badges. Importantly, none of the reproducibility issues we encountered appeared seriously consequential for the conclusions stated in the original articles. Nevertheless, non-reproducibility highlighted fundamental quality control and documentation failures during data management, data analysis, and reporting. Furthermore, the extensive time investment required to check and establish analytic reproducibility would likely be prohibitive for many researchers interested in re-using data. The findings are consistent with a previous study which found a similar rate of non-reproducibility and identified unclear, incomplete, or incorrect analysis reporting as a primary contributing factor [@hardwicke2018a]. Although the open badges scheme introduced at Psychological Science has been associated with an increase in data availability [@kidwell2016], the current findings suggest that additional efforts may be required in order to ensure analytic reproducibility.
Although some reproducibility issues were resolved through involvement of original authors, relying on such assistance is not ideal for several reasons: (1) it substantially increases the workload and time investment, both for researchers attempting to re-use data and for original authors; (2) important details about and materials related to the study are more likely to be forgotten or misplaced over time, reducing the ability of authors to assist [@vines2014]; (3) authors may be unwilling or unable to respond, as was the case for three of the articles we examined. Author involvement highlighted that non-reproducibility was often caused by unclear, incomplete, or incorrect reporting of analytic procedures in the original papers, issues which could often be resolved through the provision of additional information or clarifications beyond what was previously reported. In other cases, authors informed us of errors in shared data files or analysis scripts, suggesting that at some stage the original data, analyses, and research report had become decoupled. For many issues, neither we nor the original authors could reproduce the target values and the causal locus of non-reproducibility remains unidentified; it was no longer possible to reconstruct the original analysis pipeline.

@@

## Conclusion
This study has highlighted that the analytic reproducibility of published psychology articles cannot be guaranteed. It is inevitable that some scientific manuscripts contain errors and imperfections, as researchers are only human and people make mistakes [@rouder2019]. However, most of the issues we encountered in this study were entirely avoidable. Data availability alone is insufficient; further action is required to ensure the analytic reproducibility of scientific manuscripts.

@@

TEH and MCF designed the study. TEH, MB, KM, EH, MBN, BNP, BdM, BL, EJY, and MCF conducted reproducibility checks. TEH performed the data analysis. TEH and MCF wrote the manuscript. MBN provided feedback on the manuscript. All authors gave final approval for publication.

>>>>

author: 
  - name          : "Tom E. Hardwicke"
    affiliation   : "1,2"
    corresponding : yes    # Define only one corresponding author
    address       : "Nieuwe Achtergracht 129B, Department of Psychology, University of Amsterdam, 1018 WT Amsterdam, The Netherlands"
    email         : "tom.hardwicke@uva.nl"
  - name          : "Manuel Bohn"
    affiliation   : "3"
  - name          : "Kyle MacDonald"
    affiliation   : "4"
  - name          : "Emily Hembacher"
    affiliation   : "5"    
  - name          : "Michèle B. Nuijten"
    affiliation   : "6"
  - name          : "Benjamin N. Peloquin"
    affiliation   : "5"
  - name          : "Benjamin E. deMayo"
    affiliation   : "5"
  - name          : "Bria Long"
    affiliation   : "5"
  - name          : "Erica J. Yoon"
    affiliation   : "5"
  - name          : "Michael C. Frank"
    affiliation   : "5"

affiliation:
  - id            : "1"
    institution   : "University of Amsterdam"
  - id            : "2"
    institution   : "Charité – Universitätsmedizin Berlin"
  - id            : "3"
    institution   : "Max Planck Institute for Evolutionary Anthropology"
  - id            : "4"
    institution   : "University of California, Los Angeles"
  - id            : "5"
    institution   : "Stanford University"
  - id            : "6"
    institution   : "Tilburg University"

@@

abstract: | 
  For any scientific report, repeating the original analyses upon the original data should yield the original outcomes. We evaluated analytic reproducibility in `r knitr::load_cache(label = 'abstract_text', object = 'totalArticles')` Psychological Science articles awarded open data badges between 2014-2015. Initially, `r knitr::load_cache(label = 'abstract_text', object = 'intialReproRateArticles')` articles contained at least one “major numerical discrepancy” (>10% difference) prompting us to request input from original authors. Ultimately, target values were reproducible without author involvement for `r knitr::load_cache(label = 'abstract_text', object = 'repro_noAI')` articles; reproducible with author involvement for `r knitr::load_cache(label = 'abstract_text', object = 'repro_withAI')` articles; not fully reproducible with no substantive author response for `r knitr::load_cache(label = 'abstract_text', object = 'noRepro_noAI')` articles; and not fully reproducible despite author involvement for `r knitr::load_cache(label = 'abstract_text', object = 'noRepro_withAI')` articles. Overall, `r knitr::load_cache(label = 'abstract_text', object = 'majorErrors')` major numerical discrepancies remained out of `r knitr::load_cache(label = 'abstract_text', object = 'totalValues')` checked values `r knitr::load_cache(label = 'abstract_text', object = 'reproRateValues')`, but original conclusions did not appear affected. Non-reproducibility was primarily caused by unclear analysis reporting. These results highlight that open data alone is not sufficient to ensure analytic reproducibility.

@@

```{r load_packages}
# load packages
library(knitr) # for literate programming
library(papaja) # for article template
library(tidyverse) # for data munging
library(tidylog) # inline analysis feedback
library(here) # for finding files
library(DescTools) # for calculating confidence intervals
library(english) # for converting numbers to words
#library(metafor) # for meta-analysis
library(meta) # for forest plot
library(osfr) # to download data from the Open Science Framework (for meta-analysis)
```

```{r load_data}
# loads the processed data files output by the preprocessing performed above
load(here('data', 'processed', 'data_values.RData')) # load value-level data
load(here('data', 'processed', 'data_articles.RData')) # load article-level data
load(here('data', 'processed', 'data_kidwell.RData')) # load kidwell2016 data
load(here('data', 'processed', 'data_hardwicke2018.RData')) # load hardwicke2018 data
```

@@

@@

[^compRepro]: Computational reproducibility is often assessed by attempting to re-run original computational code and can therefore fail if original code is unavailable or non-functioning [e.g., @stodden2018]. By contrast, analytic reproducibility is assessed by attempting to repeat the original analysis procedures, which can involve implementing those procedures in new code if necessary [e.g., @hardwicke2018a].

@@

Of `r data_kidwell_badge %>% nrow()` articles marked with an open data badge, Kidwell and colleagues had identified `r data_kidwell_badge_reusable %>% nrow()` with datasets that met four reusability criteria (accessible, correct, complete, and understandable). For each of these articles, one investigator (TEH) attempted to identify a coherent set of descriptive and inferential statistics (e.g., means, standard deviations, t-values, p-values; see Figure \@ref(fig:valueTypePlot)), roughly 2-3 paragraphs of text, sometimes including a table or figure, related to a ‘substantive’ finding based on ‘relatively straightforward’ analyses. We focused on substantive findings because they are the most important and straightforward analyses to ensure that our team had sufficient expertise to re-run them. In total, `r nrow(data_values)` discrete numerical values reported in `r totalArticles` articles were identified as target values. The articles were published between January 2014 and May 2015. Further information about the sample is available in Supplementary Information \@ref(sup_sample).

@@

The present study is highly comparable in goal and design to a previous study (Hardwicke et al. 2018), presenting an opportunity to synthesize their estimates of analytic reproducibility in psychology. To this end, we used random-effects models to meta-analyze the raw proportion estimates of article-level analytic reproducibility for the two studies, separately before and after original authors were contacted.

@@

```{r}
# prepare information on insufficent information errors for reporting inline
insuffInfoErrors <- data_articles %>% filter(Insufficient_Information_Errors > 0) %>% nrow() # how many articles had at least 1 insufficient information error
insuffInfoErrors_percent <- insuffInfoErrors/totalArticles*100 # calculate percent
insuffInfoErrors_CIs <- propCI(insuffInfoErrors,totalArticles) # caluclate CIs
```
Prior to seeking original author involvement, all target outcomes in `r reproBefore %>% filter(reproOutcome == T) %>% pull(n)` out of the `r totalArticles` articles (`r reproBefore %>% filter(reproOutcome == T) %>% pull(percent)`) were reproducible, with the remaining `r reproBefore %>% filter(reproOutcome == F) %>% pull(n)` (`r reproBefore %>% filter(reproOutcome == F) %>% pull(percent)`) containing at least one major numerical discrepancy. After requesting input from original authors, several issues were resolved through the provision of additional information[^authorInfo] and ultimately all target outcomes in `r reproAfter %>% filter(reproOutcome == T) %>% pull(n)` (`r reproAfter %>% filter(reproOutcome == T) %>% pull(percent)`) articles were reproducible, with the remaining `r reproAfter %>% filter(reproOutcome == F) %>% pull(n)` (`r reproAfter %>% filter(reproOutcome == F) %>% pull(percent)`) articles containing at least one major numerical discrepancy. The full breakdown of reproducibility outcomes after author contact are shown in Table \@ref(tab:reproTab) and the results of a meta-analysis synthesizing the findings with those of a previous study [@hardwicke2018a] are displayed in Figure \@ref(fig:forestPlot). Aspects of `r as.english(insuffInfoErrors)` reproducibility checks could not be completed as there was insufficient information to proceed with the reanalysis. Availability of original analysis scripts appeared to provide some modest benefits in terms of reproducibility outcomes and time expenditure (see Supplementary Information \@ref(sup_results)). In no cases did the observed numerical discrepancies appear to be consequential for the conclusions stated in the original articles (see Supplementary Information \@ref(vignettes)).

@@

```{r scatterPlot, fig.cap="Scatterplot showing reanalysis p-values as a function of reported p-values, classified by reproducibility outcome. Axes are on a log scale. For display purposes, 41 values below 0.001 are not shown.", fig.path="figs/"}
source(here('analysis','plot_reproScatter.R')) # build plot
scatterPlot # show plot
```
```{r}
source(here('analysis','meta_analysis.R'))
```

```{r fig.cap="Forest plot showing the proportion of articles with fully reproducible target outcomes in Hardwicke et al. (2018) and the present study before and after contacting original authors for assistance with reproducibility checks. Squares represent individual study proportions (with size proportional to meta-analytic weighting). Diamonds represent summary proportions estimated by random effects (RE) models. Error bars represent 95% confidence intervals (CIs).", fig.path="figs/"}
include_graphics(here('manuscript', 'figs', 'forestPlot.png'))
```

@@

A reasonable expectation of any scientific manuscript is that repeating the original analyses upon the original data will yield the same quantitative outcomes [@bollen2015]. We have found that this standard of analytic reproducibility was frequently not met in a sample of Psychological Science articles receiving open data badges. Importantly, none of the reproducibility issues we encountered appeared seriously consequential for the conclusions stated in the original articles. Nevertheless, non-reproducibility highlighted fundamental quality control and documentation failures during data management, data analysis, and reporting. Furthermore, the extensive time investment required to check and establish analytic reproducibility would likely be prohibitive for many researchers interested in re-using data. The findings are consistent with a previous study which found a similar rate of non-reproducibility and identified unclear, incomplete, or incorrect analysis reporting as a primary contributing factor [@hardwicke2018a]. Although the open badges scheme introduced at Psychological Science has been associated with an increase in data availability [@kidwell2016], the current findings suggest that additional efforts may be required in order to ensure analytic reproducibility.

Some reproducibility issues were resolved through involvement of original authors, however, relying on such assistance is not ideal for several reasons: (1) it substantially increases the workload and time investment, both for researchers attempting to re-use data and for original authors; (2) important details about and materials related to the study are more likely to be forgotten or misplaced over time, reducing the ability of authors to assist [@vines2014]; (3) authors may be unwilling or unable to respond, as was the case for three of the articles we examined. Author involvement highlighted that non-reproducibility was often caused by unclear, incomplete, or incorrect reporting of analytic procedures in the original papers, issues which could often be resolved through the provision of additional information or clarifications beyond what was previously reported. In other cases, authors informed us of errors in shared data files or analysis scripts, suggesting that at some stage the original data, analyses, and research report had become decoupled. For many issues, neither we nor the original authors could reproduce the target values and the causal locus of non-reproducibility remained unidentified; it was no longer possible to reconstruct the original analysis pipeline.

@@

## Conclusion
Together with previous findings [@hardwicke2018a], this study has highlighted that the analytic reproducibility of published psychology articles cannot be guaranteed. It is inevitable that some scientific manuscripts contain errors and imperfections, as researchers are only human and people make mistakes [@rouder2019]. However, most of the issues we encountered in this study were entirely avoidable. Data availability alone is insufficient; further action is required to ensure the analytic reproducibility of scientific manuscripts.

@@

TEH and MCF designed the study. TEH, MB, KM, EH, MBN, BNP, BdM, BL, EJY, and MCF conducted reproducibility checks. TEH performed the data analysis. TEH and MCF wrote the manuscript. MB & MBN provided feedback on the manuscript. All authors gave final approval for publication.
